Time Series Length at Which Block Bootstrapping is Effective for Estimation of Variance

### 2022-10-06 Manuscript/Block-Bootstrap.Rmd
Manuscript Update

do1rep <- function(n, theta, sd, l, statistic) {
  x <- arima.sim(list(ar = theta), n = n, sd = sd)

@@

@@

print(root_0.6_4)
```
8 out of the 10 coverage rates are below .90.

@@

```{r} 
colnames(root_0.2_1) <- NULL
colnames(root_0.2_2) <- NULL
colnames(root_0.2_3) <- NULL
root_0.2_3
```

@@

```{r}
root_0.4 <- data.frame(rbind(t(root_0.4_1[c(1, 2, 7, 18, 17),]), t(root_0.4_1[c(1, 2, 11, 20, 19),]), t(root_0.4_2[c(1, 2, 7, 18, 17),]), t(root_0.4_2[c(1, 2, 11, 20, 19),]), t(root_0.4_3[c(1, 2, 7, 18, 17),]), t(root_0.4_3[c(1, 2, 11, 20, 19),]), t(root_0.4_4[c(1, 2, 7, 18, 17),]), t(root_0.4_4[c(1, 2, 11, 20, 19),])))
rownames(root_0.4) <- NULL
colnames(root_0.4) <- c("n", "l", 'ci', 'lb', 'ub')
root_0.4['method'] <- c('nm', 'nm', 'nm', 'nm', 'nm', 'mv', 'mv', 'mv', 'mv', 'mv', 'nm', 'nm', 'nm', 'nm', 'nm', 'mv', 'mv', 'mv', 'mv', 'mv', 'nm', 'nm', 'nm', 'nm', 'nm', 'mv', 'mv', 'mv', 'mv', 'mv', 'nm', 'nm', 'nm', 'nm', 'nm', 'mv', 'mv', 'mv', 'mv', 'mv')
root_0.4[order(-root_0.4$n), ]
```
```{r}
library(ggplot2)
ggplot(root_0.4, aes(x = n, y = ci, color = method)) +
  geom_errorbar(aes(ymin=lb, ymax=ub), colour="black", width=10) +
  geom_line() +
  geom_point() +
  facet_wrap(~method, labeller=labeller(method = c(mv = 'Moving Method', nm = 'Non-Moving Method'))) +
  labs(title = expression('l = ' * sqrt(n) * ', AR = 0.4'), x = 'Time Series Length', y = 'Confidence Interval Mean Recovery Rate', color = 'Method') +
  theme(legend.position="none") +
  scale_color_manual(labels = c("Moving Method", "Non-Moving Method"), values = c("blue", "red"))
ggsave('root_0.4.pdf')
```
```{r}
root_0.6 <- data.frame(rbind(t(root_0.6_1[c(1, 2, 7, 18, 17),]), t(root_0.6_1[c(1, 2, 11, 20, 19),]), t(root_0.6_2[c(1, 2, 7, 18, 17),]), t(root_0.6_2[c(1, 2, 11, 20, 19),]), t(root_0.6_3[c(1, 2, 7, 18, 17),]), t(root_0.6_3[c(1, 2, 11, 20, 19),]), t(root_0.6_4[c(1, 2, 7, 18, 17),]), t(root_0.6_4[c(1, 2, 11, 20, 19),])))
rownames(root_0.6) <- NULL
colnames(root_0.6) <- c("n", "l", 'ci', 'lb', 'ub')
root_0.6['method'] <- c('nm', 'nm', 'nm', 'nm', 'nm', 'mv', 'mv', 'mv', 'mv', 'mv', 'nm', 'nm', 'nm', 'nm', 'nm', 'mv', 'mv', 'mv', 'mv', 'mv', 'nm', 'nm', 'nm', 'nm', 'nm', 'mv', 'mv', 'mv', 'mv', 'mv', 'nm', 'nm', 'nm', 'nm', 'nm', 'mv', 'mv', 'mv', 'mv', 'mv')
root_0.6[order(-root_0.6$n), ]
```

```{r}
library(ggplot2)
ggplot(root_0.6, aes(x = n, y = ci, color = method)) +
  geom_errorbar(aes(ymin=lb, ymax=ub), colour="black", width=15) +
  geom_line() +
  geom_point() +
  facet_wrap(~method, labeller=labeller(method = c(mv = 'Moving Method', nm = 'Non-Moving Method'))) +
  labs(title = expression('l = ' * sqrt(n) * ', AR = 0.6'), x = 'Time Series Length', y = 'Confidence Interval Mean Recovery Rate', color = 'Method') +
  theme(legend.position="none") +
  scale_color_manual(labels = c("Moving Method", "Non-Moving Method"), values = c("blue", "red"))
ggsave('root_0.6.pdf')
```

>>>>

do1rep <- function(n, theta, sd, l, statistic) {
  x <- arima.sim(list(ar = theta), n = n, sd = sqrt((sd^2)*(1-theta^2)))

do1rep2 <- function(n, theta, sd, l, statistic) {
  x <- rnorm(n = n, sd = sd)

@@

```{r functions, echo = False}
set.seed(5)
o = my_summary(replicate(1000, do1rep2(400, 0, 0.1, 20, m1)))
root_norm_1 <- data.frame(c(400, 20, o))
for(l in 19:16)
{
  set.seed(5)
  n = l^2
  t = my_summary(replicate(1000, do1rep2(n, 0, 0.1, l, m1)))
  root_norm_1 <- cbind(root_norm_1, c(n, l, t))

print(root_norm_1)
```

```{r functions, echo = False}
set.seed(5)
o = my_summary(replicate(1000, do1rep2(225, 0, 0.1, 15, m1)))
root_norm_2 <- data.frame(c(225, 15, o))
for(l in 14:11)
{
  set.seed(5)
  n = l^2
  t = my_summary(replicate(1000, do1rep2(n, 0, 0.1, l, m1)))
  root_norm_2 <- cbind(root_norm_2, c(n, l, t))

print(root_norm_2)
```

```{r functions, echo = False}
set.seed(5)
o = my_summary(replicate(1000, do1rep2(100, 0, 0.1, 10, m1)))
root_norm_3 <- data.frame(c(100, 10, o))
for(l in 9:6)
{
  set.seed(5)
  n = l^2
  t = my_summary(replicate(1000, do1rep2(n, 0, 0.1, l, m1)))
  root_norm_3 <- cbind(root_norm_3, c(n, l, t))

print(root_norm_3)
```

```{r functions, echo = False}
set.seed(5)
o = my_summary(replicate(1000, do1rep(400, 0, 0.1, 20, m1)))
root_0_1 <- data.frame(c(400, 20, o))
for(l in 19:16)
{
  set.seed(5)
  n = l^2
  t = my_summary(replicate(1000, do1rep(n, 0, 0.1, l, m1)))
  root_0_1 <- cbind(root_0_1, c(n, l, t))

print(root_0_1)
```

```{r functions, echo = False}
set.seed(5)
o = my_summary(replicate(1000, do1rep(225, 0, 0.1, 15, m1)))
root_0_2 <- data.frame(c(225, 15, o))
for(l in 14:11)
{
  set.seed(5)
  n = l^2
  t = my_summary(replicate(1000, do1rep(n, 0, 0.1, l, m1)))
  root_0_2 <- cbind(root_0_2, c(n, l, t))

print(root_0_2)
```

```{r functions, echo = False}
set.seed(5)
o = my_summary(replicate(1000, do1rep(100, 0, 0.1, 10, m1)))
root_0_3 <- data.frame(c(100, 10, o))
for(l in 9:6)
{
  set.seed(5)
  n = l^2
  t = my_summary(replicate(1000, do1rep(n, 0, 0.1, l, m1)))
  root_0_3 <- cbind(root_0_3, c(n, l, t))

print(root_0_3)
```

@@

print(root_0.6_4)
```
8 out of the 10 coverage rates are below .90.
```{r functions, echo = False}
set.seed(5)
o = my_summary(replicate(1000, do1rep2(400, 0, 0.1, 10, m1)))
constant_norm_1 <- data.frame(c(400, 10, o))
for(m in 19:16)
{
  set.seed(5)
  n = 20 * m
  t = my_summary(replicate(1000, do1rep2(n, 0, 0.1, 10, m1)))
  constant_norm_1 <- cbind(constant_norm_1, c(n, 10, t))

print(constant_norm_1)
```
Coverage rates are good.

```{r functions, echo = False}
set.seed(5)
o = my_summary(replicate(1000, do1rep2(300, 0, 0.1, 10, m1)))
constant_norm_2 <- data.frame(c(300, 10, o))
for(m in 14:11)
{
  set.seed(5)
  n = 20 * m
  t = my_summary(replicate(1000, do1rep2(n, 0, 0.1, 10, m1)))
  constant_norm_2 <- cbind(constant_norm_2, c(n, 10, t))

print(constant_norm_2)
```
Coverage rates are still good.

```{r functions, echo = False}
set.seed(5)
o = my_summary(replicate(1000, do1rep2(200, 0, 0.1, 10, m1)))
constant_norm_3 <- data.frame(c(200, 10, o))
for(m in 9:6)
{
  set.seed(5)
  n = 20 * m
  t = my_summary(replicate(1000, do1rep2(n, 0, 0.1, 10, m1)))
  constant_norm_3 <- cbind(constant_norm_3, c(n, 10, t))

print(constant_norm_3)
```
For n <= 180, coverage rates are below .92. For n = 140, both coverage rates are below .90, and the moving method has a coverage rate below .90 for n = 120.

```{r functions, echo = False}
set.seed(5)
o = my_summary(replicate(1000, do1rep2(100, 0, 0.1, 10, m1)))
constant_norm_4 <- data.frame(c(100, 10, o))
for(m in 4:1)
{
  set.seed(5)
  n = 20 * m
  t = my_summary(replicate(1000, do1rep2(n, 0, 0.1, 10, m1)))
  constant_norm_4 <- cbind(constant_norm_4, c(n, 10, t))

print(constant_norm_4)
```

```{r functions, echo = False}
set.seed(5)
o = my_summary(replicate(1000, do1rep(400, 0, 0.1, 10, m1)))
constant_0_1 <- data.frame(c(400, 10, o))
for(m in 19:16)
{
  set.seed(5)
  n = 20 * m
  t = my_summary(replicate(1000, do1rep(n, 0, 0.1, 10, m1)))
  constant_0_1 <- cbind(constant_0_1, c(n, 10, t))

print(constant_0_1)
```
Coverage rates are good.

```{r functions, echo = False}
set.seed(5)
o = my_summary(replicate(1000, do1rep(300, 0, 0.1, 10, m1)))
constant_0_2 <- data.frame(c(300, 10, o))
for(m in 14:11)
{
  set.seed(5)
  n = 20 * m
  t = my_summary(replicate(1000, do1rep(n, 0, 0.1, 10, m1)))
  constant_0_2 <- cbind(constant_0_2, c(n, 10, t))

print(constant_0_2)
```
Coverage rates are still good.

```{r functions, echo = False}
set.seed(5)
o = my_summary(replicate(1000, do1rep(200, 0, 0.1, 10, m1)))
constant_0_3 <- data.frame(c(200, 10, o))
for(m in 9:6)
{
  set.seed(5)
  n = 20 * m
  t = my_summary(replicate(1000, do1rep(n, 0, 0.1, 10, m1)))
  constant_0_3 <- cbind(constant_0_3, c(n, 10, t))

print(constant_0_3)
```
For n <= 180, coverage rates are below .92. For n = 140, both coverage rates are below .90, and the moving method has a coverage rate below .90 for n = 120.

```{r functions, echo = False}
set.seed(5)
o = my_summary(replicate(1000, do1rep(100, 0, 0.1, 10, m1)))
constant_0_4 <- data.frame(c(100, 10, o))
for(m in 4:1)
{
  set.seed(5)
  n = 20 * m
  t = my_summary(replicate(1000, do1rep(n, 0, 0.1, 10, m1)))
  constant_0_4 <- cbind(constant_0_4, c(n, 10, t))

print(constant_0_4)
```

@@

```{r} 
colnames(root_0.2_1) <- NULL
colnames(root_0.2_2) <- NULL
colnames(root_0.2_3) <- NULL
root_0.2_3
```
```{r}

root_norm <- data.frame(rbind(t(root_norm_1[c(1, 2, 7, 18, 17),]), t(root_norm_1[c(1, 2, 11, 20, 19),]), t(root_norm_2[c(1, 2, 7, 18, 17),]), t(root_norm_2[c(1, 2, 11, 20, 19),]), t(root_norm_3[c(1, 2, 7, 18, 17),]), t(root_norm_3[c(1, 2, 11, 20, 19),])))
rownames(root_norm) <- NULL
colnames(root_norm) <- c("n", "l", 'coverage_rate', 'lb', 'ub')
root_norm['method'] <- c('nm', 'nm', 'nm', 'nm', 'nm', 'mv', 'mv', 'mv', 'mv', 'mv', 'nm', 'nm', 'nm', 'nm', 'nm', 'mv', 'mv', 'mv', 'mv', 'mv', 'nm', 'nm', 'nm', 'nm', 'nm', 'mv', 'mv', 'mv', 'mv', 'mv')
root_norm[order(-root_norm$n), ]
```

```{r}
library(ggplot2)
ggplot(root_norm, aes(x = n, y = coverage_rate, color = method)) +
  geom_errorbar(aes(ymin=lb, ymax=ub), colour="black", width=5) +
  geom_line() +
  geom_point() +
  facet_wrap(~method, labeller=labeller(method = c(mv = 'Moving Method', nm = 'Non-Moving Method'))) +
  labs(title = expression('l = ' * sqrt(n) * ', AR = 0'), x = 'Time Series Length', y = 'Confidence Interval Mean Recovery Rate', color = 'Method') +
  theme(legend.position="none") +
  scale_color_manual(labels = c("Moving Method", "Non-Moving Method"), values = c("blue", "red"))
ggsave('root_norm.pdf')
```

root_0 <- data.frame(rbind(t(root_0_1[c(1, 2, 7, 18, 17),]), t(root_0_1[c(1, 2, 11, 20, 19),]), t(root_0_2[c(1, 2, 7, 18, 17),]), t(root_0_2[c(1, 2, 11, 20, 19),]), t(root_0_3[c(1, 2, 7, 18, 17),]), t(root_0_3[c(1, 2, 11, 20, 19),])))
rownames(root_0) <- NULL
colnames(root_0) <- c("n", "l", 'coverage_rate', 'lb', 'ub')
root_0['method'] <- c('nm', 'nm', 'nm', 'nm', 'nm', 'mv', 'mv', 'mv', 'mv', 'mv', 'nm', 'nm', 'nm', 'nm', 'nm', 'mv', 'mv', 'mv', 'mv', 'mv', 'nm', 'nm', 'nm', 'nm', 'nm', 'mv', 'mv', 'mv', 'mv', 'mv')
root_0[order(-root_0$n), ]
```

```{r}
library(ggplot2)
ggplot(root_0, aes(x = n, y = coverage_rate, color = method)) +
  geom_errorbar(aes(ymin=lb, ymax=ub), colour="black", width=5) +
  geom_line() +
  geom_point() +
  facet_wrap(~method, labeller=labeller(method = c(mv = 'Moving Method', nm = 'Non-Moving Method'))) +
  labs(title = expression('l = ' * sqrt(n) * ', AR = 0'), x = 'Time Series Length', y = 'Confidence Interval Mean Recovery Rate', color = 'Method') +
  theme(legend.position="none") +
  scale_color_manual(labels = c("Moving Method", "Non-Moving Method"), values = c("blue", "red"))
ggsave('root_0.pdf')
```

@@

```{r}
root_0.4 <- data.frame(rbind(t(root_0.4_1[c(1, 2, 7, 18, 17),]), t(root_0.4_1[c(1, 2, 11, 20, 19),]), t(root_0.4_2[c(1, 2, 7, 18, 17),]), t(root_0.4_2[c(1, 2, 11, 20, 19),]), t(root_0.4_3[c(1, 2, 7, 18, 17),]), t(root_0.4_3[c(1, 2, 11, 20, 19),]), t(root_0.4_4[c(1, 2, 7, 18, 17),]), t(root_0.4_4[c(1, 2, 11, 20, 19),])))
rownames(root_0.4) <- NULL
colnames(root_0.4) <- c("n", "l", 'ci', 'lb', 'ub')
root_0.4['method'] <- c('nm', 'nm', 'nm', 'nm', 'nm', 'mv', 'mv', 'mv', 'mv', 'mv', 'nm', 'nm', 'nm', 'nm', 'nm', 'mv', 'mv', 'mv', 'mv', 'mv', 'nm', 'nm', 'nm', 'nm', 'nm', 'mv', 'mv', 'mv', 'mv', 'mv', 'nm', 'nm', 'nm', 'nm', 'nm', 'mv', 'mv', 'mv', 'mv', 'mv')
root_0.4[order(-root_0.4$n), ]
```
```{r}
library(ggplot2)
ggplot(root_0.4, aes(x = n, y = ci, color = method)) +
  geom_errorbar(aes(ymin=lb, ymax=ub), colour="black", width=10) +
  geom_line() +
  geom_point() +
  facet_wrap(~method, labeller=labeller(method = c(mv = 'Moving Method', nm = 'Non-Moving Method'))) +
  labs(title = expression('l = ' * sqrt(n) * ', AR = 0.4'), x = 'Time Series Length', y = 'Confidence Interval Mean Recovery Rate', color = 'Method') +
  theme(legend.position="none") +
  scale_color_manual(labels = c("Moving Method", "Non-Moving Method"), values = c("blue", "red"))
ggsave('root_0.4.pdf')
```

```{r}
root_0.6 <- data.frame(rbind(t(root_0.6_1[c(1, 2, 7, 18, 17),]), t(root_0.6_1[c(1, 2, 11, 20, 19),]), t(root_0.6_2[c(1, 2, 7, 18, 17),]), t(root_0.6_2[c(1, 2, 11, 20, 19),]), t(root_0.6_3[c(1, 2, 7, 18, 17),]), t(root_0.6_3[c(1, 2, 11, 20, 19),]), t(root_0.6_4[c(1, 2, 7, 18, 17),]), t(root_0.6_4[c(1, 2, 11, 20, 19),])))
rownames(root_0.6) <- NULL
colnames(root_0.6) <- c("n", "l", 'ci', 'lb', 'ub')
root_0.6['method'] <- c('nm', 'nm', 'nm', 'nm', 'nm', 'mv', 'mv', 'mv', 'mv', 'mv', 'nm', 'nm', 'nm', 'nm', 'nm', 'mv', 'mv', 'mv', 'mv', 'mv', 'nm', 'nm', 'nm', 'nm', 'nm', 'mv', 'mv', 'mv', 'mv', 'mv', 'nm', 'nm', 'nm', 'nm', 'nm', 'mv', 'mv', 'mv', 'mv', 'mv')
root_0.6[order(-root_0.6$n), ]
```

```{r}
library(ggplot2)
ggplot(root_0.6, aes(x = n, y = ci, color = method)) +
  geom_errorbar(aes(ymin=lb, ymax=ub), colour="black", width=15) +
  geom_line() +
  geom_point() +
  facet_wrap(~method, labeller=labeller(method = c(mv = 'Moving Method', nm = 'Non-Moving Method'))) +
  labs(title = expression('l = ' * sqrt(n) * ', AR = 0.6'), x = 'Time Series Length', y = 'Confidence Interval Mean Recovery Rate', color = 'Method') +
  theme(legend.position="none") +
  scale_color_manual(labels = c("Moving Method", "Non-Moving Method"), values = c("blue", "red"))
ggsave('root_0.6.pdf')
```
```{r}
constant_norm <- data.frame(rbind(t(constant_norm_1[c(1, 2, 7, 18, 17),]), t(constant_norm_1[c(1, 2, 11, 20, 19),]), t(constant_norm_2[c(1, 2, 7, 18, 17),]), t(constant_norm_2[c(1, 2, 11, 20, 19),]), t(constant_norm_3[c(1, 2, 7, 18, 17),]), t(constant_norm_3[c(1, 2, 11, 20, 19),]), t(constant_norm_4[c(1, 2, 7, 18, 17),]), t(constant_norm_4[c(1, 2, 11, 20, 19),])))
rownames(constant_norm) <- NULL
colnames(constant_norm) <- c("n", "l", 'ci', 'lb', 'ub')
constant_norm['method'] <- c('nm', 'nm', 'nm', 'nm', 'nm', 'mv', 'mv', 'mv', 'mv', 'mv', 'nm', 'nm', 'nm', 'nm', 'nm', 'mv', 'mv', 'mv', 'mv', 'mv', 'nm', 'nm', 'nm', 'nm', 'nm', 'mv', 'mv', 'mv', 'mv', 'mv', 'nm', 'nm', 'nm', 'nm', 'nm', 'mv', 'mv', 'mv', 'mv', 'mv')
constant_norm[order(-constant_norm$n), ]
```

```{r}
library(ggplot2)
ggplot(constant_norm, aes(x = n, y = ci, color = method)) +
  geom_errorbar(aes(ymin=lb, ymax=ub), colour="black", width=5) +
  geom_line() +
  geom_point() +
  facet_wrap(~method, labeller=labeller(method = c(mv = 'Moving Method', nm = 'Non-Moving Method'))) +
  labs(title = expression('l = 10, AR = 0'), x = 'Time Series Length', y = 'Confidence Interval Mean Recovery Rate', color = 'Method') +
  theme(legend.position="none") +
  scale_color_manual(labels = c("Moving Method", "Non-Moving Method"), values = c("blue", "red"))
ggsave('constant_norm.pdf')
```

```{r}
constant_0 <- data.frame(rbind(t(constant_0_1[c(1, 2, 7, 18, 17),]), t(constant_0_1[c(1, 2, 11, 20, 19),]), t(constant_0_2[c(1, 2, 7, 18, 17),]), t(constant_0_2[c(1, 2, 11, 20, 19),]), t(constant_0_3[c(1, 2, 7, 18, 17),]), t(constant_0_3[c(1, 2, 11, 20, 19),]), t(constant_0_4[c(1, 2, 7, 18, 17),]), t(constant_0_4[c(1, 2, 11, 20, 19),])))
rownames(constant_0) <- NULL
colnames(constant_0) <- c("n", "l", 'ci', 'lb', 'ub')
constant_0['method'] <- c('nm', 'nm', 'nm', 'nm', 'nm', 'mv', 'mv', 'mv', 'mv', 'mv', 'nm', 'nm', 'nm', 'nm', 'nm', 'mv', 'mv', 'mv', 'mv', 'mv', 'nm', 'nm', 'nm', 'nm', 'nm', 'mv', 'mv', 'mv', 'mv', 'mv', 'nm', 'nm', 'nm', 'nm', 'nm', 'mv', 'mv', 'mv', 'mv', 'mv')
constant_0[order(-constant_0$n), ]
```

```{r}
library(ggplot2)
ggplot(constant_0, aes(x = n, y = ci, color = method)) +
  geom_errorbar(aes(ymin=lb, ymax=ub), colour="black", width=5) +
  geom_line() +
  geom_point() +
  facet_wrap(~method, labeller=labeller(method = c(mv = 'Moving Method', nm = 'Non-Moving Method'))) +
  labs(title = expression('l = 10, AR = 0'), x = 'Time Series Length', y = 'Confidence Interval Mean Recovery Rate', color = 'Method') +
  theme(legend.position="none") +
  scale_color_manual(labels = c("Moving Method", "Non-Moving Method"), values = c("blue", "red"))
ggsave('constant_0.pdf')
```

### 2022-10-06 Manuscript/manuscript.tex
Manuscript Update

\begin{abstract}
Block bootstrapping is a method that can be used for estimating a statistic of a time
series. It involves splitting a series into blocks (in order to account for the time
factor) and re-sampling the blocks to create many new bootstrapped time series.
This method becomes more effective as the length of the time series increases. 
The question for this study is how does one determine at what length the block
bootstrap method stops being an effective method to estimate a statistic of a time
series.

@@

Block bootstrapping is a method that can be applied to time series to estimate a parameter and it's variance. Suppose that a sample of size n of a time series is given. In order to account for the temporal dependence, the time series can be split into blocks, typically of the same size. The block should be of size l large enough to include time dependence, yet small enough to include some variance. The designer of the study can choose whether the blocks overlap or not. If the bloc(Figure here) A bootstrapped sample is created by taking a sample of n / l blocks with replacement to create a new series of size n. An estimator of the parameter is computed from the sample. This procedure can be repeated many times to create a sampling distribution if the distribution is unknown. Using this sampling distribution, a confidence interval for the parameter can be created using the alpha/2 and 1 - alpha/2 percentiles.

@@

The basic objective of the study was to assess how well block bootstrap estimates the variance for an autoregressive process. In the previous section, the process for creating a confidence interval using the block bootstrap method is described. This process is applied to a simulation of an autoregressive integrated moving average (ARIMA) with sigma = 0.1 to create a confidence interval of the mean of the process. For a certain AR coefficient, a block bootstrap 95\% confidence interval is replicated 1000 times at a starting sample size at which block bootstrap is expected to be very effective. For each individual confidence interval, it can be recorded whether or not the interval includes the true mean (0). The coverage rate, or the proportion of intervals that include the mean, can then be recorded. If the variance is being properly estimated by the block bootstrap, the coverage rate should reflect this by being close to the confidence level of 95\%. If this is not the case, the method is not working as well as it should. Since the coverage rate is a proportion, a 95\% confidence interval of the coverage rate can also be created.

@@

Before each simulation, the random seed was set to 5 so that the randomized results can be 
reproduced.

A simulation of an autoregressive integrated moving average (ARIMA) process was run 
repeatedly with varying parameters. 
The mean and standard deviation of the time series was held constant at 0 and 0.1, whereas 
the sample size and AR Coefficient
 were varied. The AR Coefficients that were used were 0.2, 0.4, and 0.6. The sample sizes 
 that were used depended on the AR
 Coefficient, and the function used to compute l. For each of these variations, 1000 
 confidence intervals are created using
 the block bootstrap method, and the proportion of these intervals that recover the true 
 mean (0) is recorded as the coverage
 rate. The block bootstrap method has different settings that were varied in this study: 
 non-moving vs moving, 
and the function for block length - the two used were a constant block length of 10 and 
the square root of n.

For the simulations using a constant block length of 10, the sample size was gradually 
lowered by increments of 20 to see when
the block bootstrap would begin to consistently return low coverage rates. For AR = 0.2 
using a constant l, results for
sample sizes ranging from 400 to 20 were observed. For AR = 0.4 using a constant l, 
results for sample sizes ranging from 600
to 20 were observed. For AR = 0.6, using a constant l, results for sample sizes ranging 
from 900 to 120 were observed. For the simulations in which the block length was the 
square root of the time series length, the only sample sizes used were perfect squares so 
that the blocks would have whole number lengths. For AR = 0.2 using l = sqrt{n}, results 
for sample sizes ranging from 625 to 36 were observed. For AR = 0.4 using sqrt{n}, results 
for sample sizes ranging from 900 to 121 were observed. For AR = 0.6 using l = sqrt{n}, 
results for sample sizes ranging from 1600 to 121 were observed.

@@

@@

>>>>

\begin{abstract}
Block bootstrapping is a method that can be used for estimating a parameter of a time
series. It involves splitting a series into blocks (in order to account for the time
factor) and re-sampling the blocks to create many new bootstrapped time series.
This method becomes more effective as the length of the time series increases. 
The question for this study is how does one determine at what length the block
bootstrap method stops being an effective method to estimate a parameter of a time
series.

@@

Block bootstrapping is a method that can be applied to a time series (or any sample of data for that matter) to estimate a parameter and its variance. Suppose that a sample of size n of a time series is given. In regular bootstrap procedure, the new sample of the same size as the original would just be created by resampling observations with replacement. In the case of a time series, in order to account for the temporal dependence, the time series can be split into blocks, typically of the same size. The block should be of size l large enough to include time dependence, yet small enough to include some variance. A bootstrapped sample is created by taking a sample of n / l blocks with replacement to create a new series of size n. The designer of the study can choose whether the blocks overlap or not. If the blocks can never overlap, it is called a non-moving block bootstrap. In this case, the original sample is split evenly into n / l blocks, and the same number of blocks is sampled without replacement to create a bootstrapped sample.

\begin{figure}[h]
  \centering
  \includegraphics[width=\textwidth]{orig_ts}
  \caption{}
  \label{fig:orig_ts}
\end{figure}

\begin{figure}[h]
  \centering
  \includegraphics[width=\textwidth]{nm_ts}
  \caption{}
  \label{fig:nm_ts}
\end{figure}

If the block do overlap, it is called a moving block bootstrap. n / l samples are taken randomly from the original sample, but they do not have to be from a group of evenly spaced blocks.

\begin{figure}[h]
  \centering
  \includegraphics[width=\textwidth]{orig_ts_2}
  \caption{}
  \label{fig:orig_ts_2}
\end{figure}

\begin{figure}[h]
  \centering
  \includegraphics[width=\textwidth]{mv_ts}
  \caption{}
  \label{fig:mv_ts}
\end{figure}

An estimator of the parameter is computed from the bootstrapped sample. This procedure can be repeated many times to create a sampling distribution of the estimator.

\begin{figure}[h]
  \centering
  \includegraphics[width=\textwidth]{nm_dist}
  \caption{}
  \label{fig:nm_dist}
\end{figure}

\begin{figure}[h]
  \centering
  \includegraphics[width=\textwidth]{mv_dist}
  \caption{}
  \label{fig:mv_dist}
\end{figure}

Using this distribution of estimators, a confidence interval for the parameter can be created using the $\alpha$/2 and 1 - $\alpha$/2 percentiles.

@@

The central objective of the study was to assess at which sample size block bootstrap no longer estimates the variance well for an autoregressive process. The method will work better as sample size goes to infinity, but the goal is to see what is the smallest sample size that is acceptable. To achieve this, simulations must first be conducted with a larger sample size in order to compare to results with a smaller n. In the previous section, the procedure for creating a confidence interval using the block bootstrap method is described. This process is applied to a simulation of an autoregressive integrated moving average (ARIMA) with standard deviation = 0.1 to create a confidence interval of the mean of the process. R has a built in ARIMA simulation which allows you to set the AR coefficient and the standard deviation of the error term. Using the equation below, the sigma of the error term can be decided so that the standard deviation of Y is constant (0.1, Variance = 0.01).

\[ \sigma_{x}^{2}=\sigma_{\eta}^{2}/\left( 1-\phi^2 \right)\]
\[\sigma_{\eta}=\sqrt{\sigma_{x}^{2}*\left( 1-\phi^2 \right)}\]

For a certain AR coefficient, a block bootstrap 95\% confidence interval is replicated 1000 times at a starting sample size at which block bootstrap is expected to be very effective (a relatively large sample). For each individual confidence interval, it can be recorded whether or not the interval includes the true mean (0). The coverage rate, or the proportion of intervals that include the mean, can then be recorded. If the variance is being properly estimated by the block bootstrap, the coverage rate should reflect this by being close to the confidence level of 95\%. If this is not the case, the method is not working as well as it should. Since the coverage rate is a proportion, a 95\% confidence interval of the coverage rate can also be created. Once the simulation is done for the starting n, the sample size can gradually be lowered until the coverage rates and their respective confidence intervals indicate that the method no longer functions well.

@@

Before each simulation (1000 replications of confidence intervals from block bootstrapped sample mean distributions), the random seed was set to 5 so that the randomized results can be 
reproduced.

A simulation of an autoregressive integrated moving average (ARIMA) process was run 
repeatedly with varying parameters. 
The mean and standard deviation of the time series was held constant at 0 and 0.1, respectively, whereas 
the sample size and AR Coefficient
 were varied. The AR Coefficients that were used were 0 (or a regular sample with no temporal dependence) 0.2, 0.4, and 0.6. The sample sizes 
 that were used depended on the AR
 Coefficient, and the function used to compute l. For each of these variations, 1000 
 confidence intervals are created using
 the block bootstrap method, and the proportion of these intervals that recover the true 
 mean (0) is recorded as the coverage
 rate. The block bootstrap method has different settings that were varied in this study: 
 non-moving vs moving, 
and the function for block length - the two used were a constant block length of 10 and 
the square root of n.

For the simulations using a constant block length of 10, the sample size was gradually 
lowered by increments of 20 to see when
the block bootstrap would begin to consistently return low coverage rates. For both AR = 0 and AR = 0.2 
using a constant l, results for
sample sizes ranging from 400 to 20 were observed. For AR = 0.4 using a constant l, 
results for sample sizes ranging from 600
to 20 were observed. For AR = 0.6, using a constant l, results for sample sizes ranging 
from 900 to 120 were observed. For the simulations in which the block length was the 
square root of the time series length, the only sample sizes used were perfect squares so 
that the blocks would have whole number lengths. For AR = 0.2 using l = $\sqrt{n}$, results 
for sample sizes ranging from 625 to 36 were observed. For AR = 0.4 using $\sqrt{n}$, results 
for sample sizes ranging from 900 to 121 were observed. For AR = 0.6 using l = $\sqrt{n}$, 
results for sample sizes ranging from 1600 to 121 were observed.

@@

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{constant_0}
  \caption{}
  \label{fig:constant_0}
\end{figure}

@@

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{root_0}
  \caption{}
  \label{fig:root_0}
\end{figure}
